---
title: 'Practical Machine Learning: Excercise Quality Prediction'
output:
  html_document:
    keep_md: yes
  pdf_document: default
---

## Introduction / Executive summary

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants was used to predict the quality of technique employed.

## Loading and exploring the data set

In first part of the code, data sets is loaded and cleaned. Useless columns with near zero variance and filled with missing data are excluded from the dataset. Type of exercises empoyed are classified as numbers from 1 to 5. Also parallel computing is setup with doMC library.

```{r, echo = FALSE, message = FALSE, warning = FALSE, results = FALSE, comment = FALSE}
library(data.table)
library(ggplot2)
library(caret)
library(doMC)

registerDoMC(cores = 4)

dataPML <- fread("./Data/pml-training.csv")
dataPMLtest <-fread("./Data/pml-testing.csv")

nsv <- nearZeroVar(dataPML, saveMetrics = T)
dataPML <- dataPML[, !nsv$nzv, with = FALSE];
dataPML[, c("V1", "user_name", "cvtd_timestamp") := NULL];
dataPMLtest[, c("V1", "user_name", "cvtd_timestamp") := NULL];
dataPML[,which(colSums(is.na(dataPML)) > 0.8 * length(rownames(dataPML))) := NULL, with = FALSE];


```

After cleaning the data, several prediction models were trained - Global Boost Machine, Random Forest.

```{r, echo = FALSE, message = FALSE, warning = FALSE, results = FALSE, comment = FALSE}

boostModel <- train(classe ~ ., method = "gbm", data = dataPML, trControl = trainControl(method = "cv", number = 10, allowParallel = TRUE))
randomForestModel <- train(classe ~ ., method = "rf", data = dataPML, importance = T, trControl = trainControl(method = "cv", number = 10, allowParallel = TRUE))


```
```{r}
boostModel
randomForestModel

```

As we can see, both of the models do very well. They achieve almost perfect predictions with accuracy of ~ 0.99. It means they almost always correctly classify quality of excercize. Training uses cross validation and chooses the model which generalizes best on the validation set. Random Forest did marginally better and is chosen for submission. 

After choosing the right model, it is used on testing set and is submitted to Coursera for grading:

```{r, echo = FALSE}

prediction <- predict(randomForestModel, dataPMLtest)

pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("./TestPredictions/problem_id_", i, ".txt")
    write.table(x[i], file = filename, quote = FALSE, row.names = FALSE, col.names = FALSE)
  }
}
pml_write_files(prediction)
```

